---
title: "Milestone 6"
author: "Irum Mehdi,Ruben Prado,Eunice Ndzerem Shang"
date: "12/14/2020"
output: html_document
urlcolor: blue
---



# Project Aim   

There is a continued need to improve immunization coverage rates.  Recent polls of adults suggest that as few as 50% of US adults are committed to receiving a COVID-19 vaccine, and misinformation and conspiracy theories about a vaccine abound. Prematurely approving a vaccine could undermine COVID-19 vaccination efforts and erode confidence in vaccines more generally (Bauchner et al., 2020). There is great concern given the flurry of public announcements from  sources  (such as the FDA, CDC, Pfizer, The White House etc.)  that have made several announcements  on soon to be vaccine COVID-19 availability. Alongside Operation Warp Speed, these announcements have  left  clinicians, public health  officials, and members of the public bewildered about what is happening and concerned that vaccines will be made available before safety and effectiveness are fully established.

In this project we set out to understand population beliefs  about vaccines in general and the drivers that influence their decisions. We aimed at answering the following questions: How do knowledge and beliefs about vaccines vary by age,race SES, income,education, location. Does it matter who is telling me about vaccines (e.g., doctors vs internet. How does insurance coverage, type of doctor,predict vaccine acceptance.What are the characteristics of vaccine hesitant groups.

## Methods
We examined two data sets

The “Anti-Vax” data set was retrieved from the University of Michigan’s repository. It is a survey summary and quantitative report on vaccine beliefs and knowledge across social media. The cross-sectional research explored the relationship between the spread of information regarding vaccines and social media use.  A sample of 2515 people over the age of 18 around the world completed the survey via a link found on Twitter, Facebook and Instagram. Data to the enormity of the data set, data cleaning and analysis for this project focused on  the United States within the North America Continent.  The methods used in data cleaning included renaming column names that were lengthy column names, recoding lengthy answers and transforming survey results  to data.frame, string normalization and coercing data to correct types (i.e. changing character variables to factors), character manipulation,  selecting specific columns  needed to answer the focus questions, finding all missing data, removing missing information by dropping unwanted values.


The Parental Attitudes data was  retrieved from PLOS Journal Research Article Titled “Parental attitudes and perceptions associated with childhood vaccine exemptions in high-exemption schools” (Pottinger et al.,2018).The survey was conducted in Arizona’s counties  with high exemption schools with the support of the Arizona Department of Public Health. The Survey was conducted in the 2012-2013 school year and a total of 405 responses were received.The study was published online in June 2018 and the associated data  and survey questionnaire was deemed reusable under the Creative Commons Attribution License.
The data set was in xlsx format . The rows with  survey questions were removed as the column headers were self explanatory and did not require name change for all the columns. Identifying variables like school names were removed to protect privacy. A single column was created for Race as it was split in various columns.Variables taking stock of exemption status were recorded from 1,0 to Yes and No.Variables such as Trust Info and Natural Immunity were based on a scale from 1-5, where 1 is ‘strongly disagree’ and 5 is ‘strongly agree’. The serious side effect variable is on the same scale going from 1-5, instead 1 is ‘not at all concerned’ and 5 is ‘extremely concerned’.For exploring the variables like concerns about vaccine safety,side effects and vaccines ability to prevent serious diseases, we changed the  column names and changed to text describing the range of range of the Likert  scale. For the exempt reason question 305 participants declined to answer. For the visualizations, counts ,proportions and percentages were calculated from variables of interest.




## Links to the data sets: 

+ [https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0198655] 


+ [https://www.openicpsr.org/openicpsr/project/120505/version/V1/view?path=/openicpsr/120505/fcr:versions/V1/Survey-Summary.xlsx&type=file]



```{r,message=FALSE,echo=FALSE,strip.white=TRUE}
library(readxl)
library(tidyverse)
library(ggthemes)
library(ggplot2)
library(ggpubr)
library(plotly)
library(lubridate)
library(kableExtra)
library(DT)

```

\newpage

```{r,echo=FALSE,message=FALSE,warning=FALSE,strip.white=TRUE}
file_path <- ("./parent.xlsx")
df <-readxl::read_excel(file_path) 
df<-slice(df,2:n())
df<-select(df,(1:39))
```



```{r,echo=FALSE,message=FALSE,warning=FALSE,strip.white=TRUE}
getRace <- function(white,asian,black,hispanic,aian,other)
{
  raceCount<-0
  race<-"Other"
  if(!is.na(white))
  {
    raceCount <- raceCount + 1
    race <- "White"
  }
  
  if(!is.na(asian))
  {
    raceCount = raceCount + 1
    race <-"Asian"
  }
  if(!is.na(black))
  {
    raceCount = raceCount + 1
    race <- "Black"
  }  
    if(!is.na(hispanic))
  {
    raceCount <- raceCount + 1
    race <- "Hispanic"
  }
  
  if(!is.na(aian))
  {
    raceCount = raceCount + 1
    race <-"Aian"
  }
  if(!is.na(other))
  {
    raceCount = raceCount + 1
    race <- "Other"
  } 
  if(raceCount > 1)
  {
    race <- "Mixed"
  }
  
  race
}

getRaceV<-Vectorize(getRace)
```




```{r,echo=FALSE,message=FALSE,warning=FALSE,strip.white=TRUE}

df_1<-df %>% 
  mutate(Race=getRaceV(White,Asian,Black,Hispanic,AIAN,Other))
race_table<-table(df_1$Race)

 df_2<- df_1 %>%
  mutate(Insurance = case_when(
          Insurance=="Yes, private insurance"~"Private",
            Insurance=="Yes, public insurance"~"Public",
          Insurance=="No"~"No Insurance",
          TRUE~NA_character_))
 # Variables of Interest  
   df_3<-df_2 %>% 
     select(Race,EverExempt,DocType,Insurance,Gender) 
     
df_5 <- df_3 %>% filter(!is.na(EverExempt))
df_6 <- df_5 %>%  mutate(Exempt = ifelse(EverExempt==1,"Yes","No"))

```






\newpage

These visualizations are from  the data from a survey of parents in high exemption schools in Arizona.Here we looked at race,gender,type of doctor,and type of insurance among survey takers to see if there are any differences among exemptors and non exemptors.

\newpage

### Type of Insurance as Predictor of Vaccine Exemption
Among parents who chose a non medical exemption in this survey,respondents with no insurance requested a non medical exemption with higher frequency.There are no significant differences  in relative frequencies between those holding Private and Public Insurance.

```{r,echo=FALSE,message=FALSE,warning=FALSE,strip.white=TRUE}

 df_7a<-df_6 %>%
  select(Exempt,Insurance) %>% 
  drop_na()

df_7b<-df_7a %>% 
  group_by(Insurance, Exempt) %>% 
  summarise(n=n()) %>% 
  mutate(pct=n/sum(n), lbl=scales::percent(pct))


b<-ggplot(df_7b, aes(x = Insurance ,
                     y = pct,
                     fill = factor(Exempt, 
                         levels = c("Yes", "No"),
                         labels = c("Yes", "No")
                         )
                     )
          ) 

b<-b+geom_bar(stat = "identity", position = "fill") 

b<-b+labs(y = "Percent", fill = "Exemption",
       x = "Insurance",
       title = "Exemption By Insurance")
b<-b+scale_y_continuous(labels = scales::percent)
b<-b+geom_text(aes(label = lbl), 
            size = 3, 
            position = position_stack(vjust = 0.5))
b<-b+scale_fill_brewer(palette = "Set2")


```
 
\newpage
### Race as Predictor of Vaccine Exemption  

Among the respondents in this survey,people of all races were similar in requesting or not requesting a non medical exemption.
There are no significant differences in relative frequencies between races.People who identify as other race were significant minority among the total respondents.

```{r,echo=FALSE,message=FALSE,warning=FALSE,strip.white=TRUE}


 df_8a<-df_6 %>%
  select(Exempt,Race) %>% 
  drop_na()

df_8b<-df_8a %>% 
  group_by(Race, Exempt) %>% 
  summarise(n=n()) %>% 
  mutate(pct=n/sum(n), lbl=scales::percent(pct))


c<-ggplot(df_8b, aes(x = Race ,
                     y = pct,
                     fill = factor(Exempt, 
                         levels = c("Yes", "No"),
                         labels = c("Yes", "No")
                         )
                     )
          ) 

c<-c+geom_bar(stat = "identity", position = "fill") 

c<-c+labs(y = "Percent", fill = "Exemption",
       x = "Race",
       title = "Exemption By Race")
c<-c+scale_y_continuous(labels = scales::percent)
c<-c+geom_text(aes(label = lbl), 
            size = 3, 
            position = position_stack(vjust = 0.5))
c<-c+scale_fill_brewer(palette = "Set2")

```

\newpage

### Type of Doctor as Predictor of Vaccine Exemption

Among the respondents  in this survey ,people without a doctor and those who visited a Naturopath requested a non medical exemption more frequently.

```{r,echo=FALSE,message=FALSE,warning=FALSE,strip.white=TRUE}

 df_9a<-df_6 %>%
  select(Exempt,DocType) %>% 
  drop_na()

df_9b<-df_9a %>% 
  group_by(DocType,Exempt) %>% 
  summarise(n=n()) %>% 
  mutate(pct=n/sum(n), lbl=scales::percent(pct))



d<-ggplot(df_9b, aes(x = DocType ,
                     y = pct,
                     fill = factor(Exempt, 
                         levels = c("Yes", "No"),
                         labels = c("Yes", "No")
                         )
                     )
          ) 

d<-d+geom_bar(stat = "identity", position = "fill") 

d<-d+labs(y = "Percent", fill = "Exemption",
       x = "Doctor Type",
       title = "Exemption By Type of Doctor")
d<-d+scale_y_continuous(labels = scales::percent)
d<-d+geom_text(aes(label = lbl), 
            size = 3, 
            position = position_stack(vjust = 0.5))
d<-d+scale_fill_brewer(palette = "Set2")


```


### GenderAs Predictor

There were no significant differences between the genders surveyed with regards to requesting a non medical exemption in this group of survey respondents.

```{r,echo=FALSE,warning=FALSE,message=FALSE}
df_9c<-df_6 %>% 
  select(Exempt,Gender) %>% 
  drop_na()


df_9d<-df_9c %>% 
  group_by(Gender,Exempt) %>% 
  summarise(n=n()) %>% 
  mutate(pct=n/sum(n), lbl=scales::percent(pct))



e<-ggplot(df_9d, aes(x = Gender ,
                     y = pct,
                     fill = factor(Exempt, 
                         levels = c("Yes", "No"),
                         labels = c("Yes", "No")
                         )
                     )
          ) 

e<-e+geom_bar(stat = "identity", position = "fill") 

e<-e+labs(y = "Percent", fill = "Exemption",
       x = "Gender",
       title = "Exemption By Gender")
e<-e+scale_y_continuous(labels = scales::percent)
e<-e+geom_text(aes(label = lbl), 
            size = 3, 
            position = position_stack(vjust = 0.5))
e<-e+scale_fill_brewer(palette = "Set2")




```



\newpage
#### Combined Plots 

```{r,echo=FALSE,message=FALSE,warning=FALSE,strip.white=TRUE,fig.height=15}
ggarrange(b,c,d,e,labels=c("B","C","D","E"), ncol=1, nrow=4)
```
\newpage

#### Respondent Characteristics by Vaccine Exemption Status

Below is a table of respondent characteristics in the table form.We have highlighted the exemptors in yellow.

```{r,echo=FALSE,message=FALSE,warning=FALSE,strip.white=TRUE}

library(DT)
charc_vacc<- df_6 %>%
  select(Exempt,Gender,Race,DocType,Insurance) %>%
  drop_na() 

datatable(charc_vacc,
          options = list(
            pageLength=5,
            columnDefs=list(
              list(className='dt-center',targets=1:3)
            ),
            dom = 'ltip'
          ),
          rownames=FALSE,
          colnames=c("Exemption Status","Gender","Race","Type of Doctor","Insurance Status"),
          filter=c("bottom"),
          editable = F,
          caption=" Respondent Charcteristics By Vaccine Exemption Status ") %>% 
          formatStyle(1,color="blue",backgroundColor = styleEqual(c("Yes","No"),c('yellow','white'))) 

  

```

\newpage
####  Reasons for Exemption 
The most common reason cited by the respondents was a lack of belief in vaccines.


```{r,echo=FALSE,figh.width=10,message=FALSE}
df_10a<-df_2 %>%
 select(-c(25:31)) 
 
df_10b <- df_10a %>% 
    filter(!is.na(EverExempt))%>% 
  mutate(EverExempt = ifelse(EverExempt==1,"Yes","No")) %>% 
  mutate(ExemptReason=case_when(
    ExemptReason=="I do not believe my child should get shots"~"Not believe in shots"  ,
    ExemptReason=="I lost my child’s shot records"~"Lost Record",
    ExemptReason=="I did not have time to go to the doctor to get updated shots"~"Time Constraint",
  ))


  
df_10c<-df_10b %>% 
  select(ExemptReason,EverExempt) %>% 
  drop_na() %>% 
  group_by(ExemptReason,EverExempt) %>% 
  summarise(n=n()) %>% 
  mutate(pct=n/sum(n), lbl=scales::percent(pct))

  
  
e<-ggplot(df_10c, aes(x = ExemptReason ,
                     y = pct,
                     fill = factor(EverExempt, 
                         levels = c("Yes", "No"),
                         labels = c("Yes", "No")
                         )
                     )
          ) 

e<-e+geom_col(position = "fill") 

e<-e+labs(y = "Percent", fill = "Exemption",
       x = "ExemptReason",
       title = "Exemption Reasons")
e<-e+scale_y_continuous(labels = scales::percent)
e<-e+geom_text(aes(label = lbl), 
            size = 3, 
            position = position_stack(vjust = 0.5))
e<-e+scale_fill_brewer(palette = "Set2")

e
```



```{r,echo=FALSE,figh.width=10,message=FALSE}
#Table of Exemptors and Non Exemptors beliefs

df_10d <- df_10a %>% 
  mutate(EverDelay =case_when(EverDelay=="1"~"Yes",
                              EverDelay=="0"~"No",
                              EverDelay=="9"~"Dont Know")) %>% 
  mutate(EverNotGet=case_when(EverNotGet==1~"Yes",
                              EverNotGet==0~"No",
                              EverNotGet==9~"Dont Know")) %>% 
  mutate(NewAllShots=case_when(NewAllShots==1~"Yes",
                               NewAllShots==0~"No",
                               NewAllShots==9~"Dont Know") ) %>%       
mutate(TooManyShots=case_when(
  TooManyShots==1~"Strongly Disagree",
  TooManyShots==2~"Disagree",
   TooManyShots==3~"Not sure",
    TooManyShots==4~"Agree",
     TooManyShots==5~"Strongly Agree"
  )) %>% 
  mutate(PreventSevereIllness=case_when(
   IllnessSevere==1~"Strongly Disagree",
   IllnessSevere==2~"Disagree",
    IllnessSevere==3~"Not sure",
     IllnessSevere==4~"Agree",
      IllnessSevere==5~"Strongly Agree",
   TRUE~NA_character_
   )) %>% 
   mutate(NaturalImmunity=case_when(
   NaturalImmunity==1~"Strongly Disagree",
   NaturalImmunity==2~"Disagree",
    NaturalImmunity==3~"Not sure",
     NaturalImmunity==4~"Agree",
      NaturalImmunity==5~"Strongly Agree",
   TRUE~NA_character_
   )) %>% 
  mutate(FewerShotsSametime=case_when(
   FewerShots==1~"Strongly Disagree",
   FewerShots==2~"Disagree",
    FewerShots==3~"Not sure",
     FewerShots==4~"Agree",
      FewerShots==5~"Strongly Agree",
   TRUE~NA_character_
   ))%>% 
  mutate(TrustInfoAboutShots=case_when(
   TrustInfo==1~"Strongly Disagree",
   TrustInfo==2~"Disagree",
    TrustInfo==3~"Not sure",
     TrustInfo==4~"Agree",
      TrustInfo==5~"Strongly Agree",
   TRUE~NA_character_
   ))%>% 
  mutate(DiscussConcerns=case_when(
   DiscussDoc==1~"Strongly Disagree",
    DiscussDoc==2~"Disagree",
     DiscussDoc==3~"Not sure",
      DiscussDoc==4~"Agree",
       DiscussDoc==5~"Strongly Agree",
   TRUE~NA_character_
   ))
```


\newpage
### Vaccine Beliefs Among Survey Takers

```{r,echo=FALSE,figh.width=10,message=FALSE}

#Vaccine Beliefs

df_10d<-df_10d %>% 
  mutate(SeriousSideEffectConcern=case_when(
  SeriousSE==1~"Not at all Concerend",
  SeriousSE==2~"Not Concerned",
   SeriousSE==3~"Not Sure",
    SeriousSE==4~"Some what Concerned",
     SeriousSE==5~"Extremely Concerned",
  TRUE~NA_character_
  )) %>% 
  mutate(ShotsNotSafeConcern=case_when(
  NotSafe==1~"Not at all Concerend",
   NotSafe==2~"Not Concerned",
    NotSafe==3~"Not Sure",
     NotSafe==4~"Some what Concerned",
      NotSafe==5~"Extremely Concerned",
  TRUE~NA_character_
  )) %>% 
  mutate(ShotsNotPreventConcern=case_when(
  NotPrevent==1~"Not at all Concerend",
    NotPrevent==2~"Not Concerned",
     NotPrevent==3~"Not Sure",
      NotPrevent==4~"Some what Concerned",
       NotPrevent==5~"Extremely Concerned",
  TRUE~NA_character_
  )) %>% 
  mutate(EverExempt = ifelse(EverExempt==1,"Yes","No"))

df_10e<-df_10d  %>% 
  select(EverExempt,SeriousSideEffectConcern,ShotsNotSafeConcern,ShotsNotPreventConcern) %>% 
  drop_na() 


df_10f<-df_10e %>% 
  group_by(EverExempt,ShotsNotSafeConcern) %>%
  tally() %>% 
mutate(percent = round(n / sum(n) * 100,1)) %>%
	dplyr::select(-n) %>%
	group_by(EverExempt) %>%
 spread(ShotsNotSafeConcern, percent)

df_10f_2 <- df_10f[,c(1,3,4,5,6,2)]
options(knitr.kable.NA = "--")
k1<-kable(df_10f_2, 
      booktabs=T,
      align='lccccc', 
      caption="Concern Regarding Vaccine Safety Percentage") %>% 
  kable_styling(full_width = F)





#Concerns regarding Side Effects

df_10g<-df_10e %>% 
  group_by(EverExempt,SeriousSideEffectConcern) %>%
  tally() %>% 
mutate(percent = round(n / sum(n) * 100,1)) %>%
	dplyr::select(-n) %>%
	group_by(EverExempt) %>%
	spread(SeriousSideEffectConcern, percent) 
      
df_10g_2 <- df_10g[,c(1,3,4,5,6,2)]
k2<-kable(df_10g_2, 
      booktabs=T,
      align='lccccc', 
      caption="Concern Regarding Vaccine Side Effects Percentage") %>% 
  kable_styling(full_width = F)

#Concerns regarding Vaccine Prevention Capability

df_10h<-df_10e %>% 
  group_by(EverExempt,ShotsNotPreventConcern) %>%
  tally() %>% 
mutate(percent = round(n / sum(n) * 100,1)) %>%
	dplyr::select(-n) %>%
	group_by(EverExempt) %>%
	spread(ShotsNotPreventConcern, percent) 
df_10h_2 <- df_10h[,c(1,3,4,5,6,2)]
k3<-kable(df_10h_2, 
      booktabs=T,
      align='lccccc', 
      caption="Concern Regarding Vaccine Preventing serious Diseases Percentage") %>% 
  kable_styling(full_width = F)
      


```

```{r echo=FALSE}
k1  

k2
k3
```

  


  


	

\newpage
#### SES,Age,Source of Information,Social Media in Vaccine Acceptance
 

```{r,message=FALSE}
Survey_Summary<-read_excel(("Survey_Summary.xlsx"))

Survey_Summary <-slice(Survey_Summary,2:n())

vac_bev_sm <- Survey_Summary %>%
  select(6, 7, 8, 9, 10, 11,13,15, 16)

vac_bev_sm <- rename(vac_bev_sm,
Age = "What is your age?",Gender = "What is your gender?", 
Continent = "Which continent do you live on?",
Education = "What is the highest level of education you have completed?", 
SES = "What socioeconomic class would you consider yourself/ your family?", 
SM_preference = "Which form of social media do you use most?", 
SM_Hrs_day = "In a typical day, how many hours do you spend on social media?",
Influene_ofSM_Info = "Information about vaccines seen on social media has:",
Most_trusted_with_Immunization_Info =
  "Who do you trust most with immunization related information/ decisions?")

vac_bev_USA <- vac_bev_sm %>% 
  filter(Continent == "North America, United States"|
           Continent == "North America, Other") %>%
   mutate(Continent = case_when(
     Continent %in% c("North America, United States") ~ "NA USA",
     Continent %in% c("North America, Other") ~ "NA Other"))
```

\newpage

```{r, message=FALSE, warning=FALSE,strip.white=TRUE}

#table(vac_bev_USA$Education)
vac_bev_USA <- vac_bev_USA %>%
  mutate(Education = case_when(
    Education %in% c("Associates Degree (2 year college/university degree)") ~ "AS",
    Education %in% c("Elementary school (grade level 1-8)") ~ "Elem",
    Education %in% c("Master's Degree") ~ "MD",
    Education %in% c("Professional/Doctoral Degree (PhD, MD, DC etc..") ~ "DD",
    Education %in% c("Bachelor Degree (4 year college/university degree)") ~ "BD",
    Education %in% c("High school (grade level 9-12/13)") ~ "HS",
    Education %in% c("No formal schooling") ~ "None",
    TRUE ~ NA_character_))

#table(vac_bev_USA$Influene_ofSM_Info)

vac_bev_USA <- vac_bev_USA %>%
  mutate(Influene_ofSM_Info = case_when(
    Influene_ofSM_Info %in% c("I have not seen anything about vaccines on social media") ~ "No Info seen",
    Influene_ofSM_Info %in% c("Influenced my opinion on vaccines (vaccines are better then I thought previously)") ~ "Positive",
    Influene_ofSM_Info %in% c("Influenced my opinion on vaccines (vaccines are worse then I thought previously)") ~ "Negative",
    Influene_ofSM_Info %in% c("NOT influenced my opinion on vaccines") ~ "None"))

#table(vac_bev_USA$SES)
vac_bev_USA$Age <-as.factor(vac_bev_USA$Age)
```

\newpage
#### Vaccine Beliefs
By Education,SES,Social Media Influence Across Different Age Groups  

```{r, message=FALSE, warning=FALSE}
# How do knowledge and beliefs about vaccines vary by age, SES, education, location:
 
vac_tab <- vac_bev_USA %>%
  mutate(count= 1) %>%
  select( Age,Education, SES,Influene_ofSM_Info, Most_trusted_with_Immunization_Info, count ) %>%
  drop_na() %>%
  group_by(Age,Education,SES,Influene_ofSM_Info) %>%
  summarize( Count = n() ) 
  data.frame()
  
  vac_bev_tablewide <- vac_tab %>%
  pivot_wider(names_from = c(Age), values_from = Count) %>%
  drop_na()
 # Created a formatted table 
  opts<-options(knitr.kable.NA="-")
kable( vac_bev_tablewide,longtable=T,booktabs=T, col.names=c("Education Level","Social Economic Status", "Influence of SM","18-24","25-34", "35-44","45-54","55-64","65+")) %>%
  #caption=" The “Anti-Vax” Movement - Social Media (SM) Influence on Knowledge and Trends of Vaccinations - North America, 2017-2018") %>%
  kable_styling(full_width=F) %>%
  kable_styling(position="left") %>%
  kable_styling(font_size=10) %>%
  add_header_above(c(" "=3,"Age Group Surveyed" =6))%>%
  column_spec(1:3, bold=TRUE) %>%
  column_spec(1:9, width="1cm") %>%
  column_spec(3:9, color= if_else(vac_bev_tablewide$Influene_ofSM_Info=="Positive", "Red","green", "green"))%>%
  pack_rows("Associate Degree",1,3)%>%
  pack_rows("Bachelors Degree",4,7)%>%
  pack_rows("High School Degree",8,10)%>%
  pack_rows("Masters Degree",11,12)%>%
  collapse_rows(columns=1, valign="top") %>%
  footnote(
  general="Benoit, Staci L. The “Anti-Vax” Movement: A Quantitative Report on Vaccine Beliefs and Knowledge across Social Media¿: Survey Summary.xlsx. Ann Arbor, MI: Inter-university Consortium for Political and Social Research [distributor], 2020-08-03. https://doi.org/10.3886/E120505V1-44300. ",
  general_title="Reference and Data Source ")

 
```

#### Proportions by Gender Surveyed
```{r,message=FALSE,warning=FALSE}
#Proportion of gender: count per category
vac_bev_plot <- vac_bev_USA %>%
  mutate(count= 1) %>%
  select( Gender, count ) %>%
  drop_na() %>%
  group_by(Gender) %>%
  summarize( num = n() ) 

vac_bev_plot$total <- sum( vac_bev_plot$num)
vac_bev_plot$prop <- vac_bev_plot$num / vac_bev_plot$total
#sum ( vac_bev_plot$prop ) 

vac_bev_plot <- vac_bev_plot %>%
  mutate(Gender = factor(Gender,
                         levels = Gender[order( num)]))

vac_bev_plot$perc <- as.character(round(vac_bev_plot$prop, 3 ) * 100 )
vac_bev_plot$lab <- paste(vac_bev_plot$num,
                          paste( " (", vac_bev_plot$perc, "%", sep = ""),
                          sep = "", ")")

ggplot(vac_bev_plot, aes( x = Gender, y = prop ) ) +
  geom_bar( stat = "identity", width = 0.3 ) +
  coord_flip() +
  ylim( 0 , 1 ) +
  geom_text( aes( x = Gender, y = prop, label = lab ),
             size = 3.5, vjust = .5, hjust = -.3 ) +
  labs(x = "Gender",
       y = "Proportion",
       title = "Gender Proportion",
       subtitle = "Proprotion by Gender surveyed") +
  theme_solarized()

```

\newpage
#### Social Media Influence on Vaccination

People between the ages of 25 and 54 and with Associate’s degree and higher were more likely to be influenced by social media platforms like Facebook. 

```{r, message=FALSE, warning=FALSE}

vac_bev_plot2 <- vac_bev_USA %>%
  mutate(count= 1) %>%
  select( Age, SM_preference, count ) %>%
  drop_na() %>%
  group_by(Age, SM_preference) %>%
  summarize( num = n() ) 

vac_bev_plot2$total <- sum( vac_bev_plot2$num)
vac_bev_plot2$prop <- vac_bev_plot2$num / vac_bev_plot2$total
#sum ( vac_bev_plot2$prop )


plot_ly(
  vac_bev_plot2,
  x= ~SM_preference,
  y= ~prop,
  color= ~Age,
  type="bar"
) %>%
  layout(title="Vaccination Trends by Scocial Media Influence, 2001-2018",
    yaxis=list(title="Proportion by Age"),
    xaxis=list(title="Social Media Preference"),
    paper_bgcolor="azure",
    plot_bgcolor="white"
  )

```

#### Resources for Vaccination Information

Doctors are the most trusted source of information regarding Vaccine Information

```{r,message=FALSE,warning=FALSE}
#Does it matter who is telling me about vaccines (e.g., doctors vs internet)?
vac_bev_plot3 <- vac_bev_USA %>%
  mutate(count= 1) %>%
  select(Influene_ofSM_Info,Most_trusted_with_Immunization_Info, count ) %>%
  drop_na() %>%
  group_by(Influene_ofSM_Info, Most_trusted_with_Immunization_Info) %>%
  summarize( Count = n() )


vac_bev_plot3 <- vac_bev_plot3 %>%
  mutate(Most_trusted_with_Immunization_Info = factor(Most_trusted_with_Immunization_Info,
                         levels = Most_trusted_with_Immunization_Info[order( Count)]))

vac_bev_plot3$total <- sum( vac_bev_plot3$Count)
vac_bev_plot3$prop <- vac_bev_plot3$Count / vac_bev_plot3$total

vac_bev_plot3$perc <- as.character(round(vac_bev_plot3$prop, 3 ) * 100 )
  vac_bev_plot3$lab <- paste(vac_bev_plot3$Count,
                          paste( " (", vac_bev_plot3$perc, "%", sep = ""),
                          sep = "", ")")

 plot_ly(
   vac_bev_plot3,
   x= ~Most_trusted_with_Immunization_Info,
  y= ~prop,
  color= ~Most_trusted_with_Immunization_Info,
  type="bar"
) %>%
  layout(barmode="stack",
           title="Resources most trusted with Vaccination Information",
    yaxis=list(title="Proportion"),
    xaxis=list(title=""),
    paper_bgcolor="azure",
    plot_bgcolor="white"
  )

```

#### Need for Immunization 

Most of the survey participants understood that vaccines are needed to achieve herd immunity to protect people who cannot be vaccinated

```{r,message=FALSE,warning=FALSE}
vac_bev_all <- Survey_Summary %>%
  select(7, 18, 19, 21, 22, 23, 26,28)

 vac_bev_all <- rename(vac_bev_all,
Gender = "What is your gender?",
RareDz_No_vac_req = "Most diseases are so rare that they no longer require vaccination.",
Herd_Immunity_needed = "Immunization of large portions of the population is needed in order to protect immunocompromised individuals (people who cannot receive vaccines).")

 vac_bev_plot28 <- vac_bev_all %>%
  mutate(count= 1) %>%
  select(Gender, RareDz_No_vac_req,Herd_Immunity_needed,count ) %>%
  drop_na()%>%
  group_by( Gender, Herd_Immunity_needed) %>%
  summarize( Count = n() )

vac_bev_plot18 <- vac_bev_all %>%
  mutate(count= 1) %>%
  select(RareDz_No_vac_req,Herd_Immunity_needed,count ) %>%
  drop_na()%>%
  group_by(RareDz_No_vac_req) %>%
  summarize( Count = n() )

ggplot( vac_bev_plot28, aes(x=Herd_Immunity_needed,y=Count)) +
  geom_bar(aes(fill="Count"), stat="identity")+
  scale_y_continuous(labels = function(x) format(x,big.mark=",",scientific=FALSE)) +
  scale_fill_manual(name="Herd Immunity Needed",values=c("coral","cyan2")) +
  # scale_fill_manual(name="Answers",values=rainbow(2))
  labs(x="",y="Population Count",title="Need to Immunize large populations",caption=" Protects people who cannot receive vaccines.") +
  theme(legend.position="bottom")


```
\newpage


#### Does Income Status Predict Vaccine Hesitany

Vaccine hesistancy does not depend on income status. Vaccine hesitancy looks to be evenly distrubuted across the different Income levels.


```{r,message=FALSE,warning=FALSE}
vaccine_new <- select(df_1, c(TrustInfo, SeriousSE, NaturalImmunity,Hesitant, ExemptReason, Income, Educ))

#vaccine_new <- vaccine_new[-1,]

```
```{r,fig.width=15}

ses<- vaccine_new %>%
  drop_na(Hesitant, Income)%>%
  group_by(Hesitant, Income) %>%
  summarise(Total= n())


ses_graph<- ggplot(ses, aes(x=Hesitant,
                            y=Total,
                            fill= Income))
ses_graph<-ses_graph+geom_bar(stat = "identity") 
ses_graph <- ses_graph + labs(x= "Vaccine Hesitance", title= "Vaccine Hesitancy and Income")

ses_graph
```

#### SocioEconomic Characteristics with Hesitancy and Exempt Reasons
```{r,message=FALSE,warning=FALSE}
ses_table <- vaccine_new %>%
  select(Hesitant,Income, Educ,ExemptReason) %>%
  drop_na() 

datatable(ses_table,
          options = list(
            columnDefs=list(
              list(className='dt-center')
            )
          ),
          rownames=FALSE,
          colnames=c("Hesitant","Income","Education","Exempt Reason"),
          caption=" Social Economic Status Characteristics with Hesitancy and Exempt Reasons")

```


\newpage

#### Summary of Results

* Insurance status and race did not significantly affect vaccine acceptance among people who sought non medical exemptions
* People who sought non medical exemptions for their children were most likely to seek health care from Naturopaths.
* Most of the survey participants understood that vaccines are needed to achieve herd immunity to protect people who cannot be vaccinated
* Doctors are the most trusted source of information regarding Vaccine Information
* People between the ages of 25 and 54 and with Associate’s degree and higher were more likely to be influenced by social media platforms like Facebook. 
 *  Middle income socioeconomic groups are more likely to be influenced by social media.
* Vaccine hesitancy does not depend on income status. 
* Vaccine hesitancy looks to be evenly distributed across the hesitant levels.


\newpage

#### Discussion

Across the different levels of vaccine hesitancy, income does not seem to be a factor. Income looks to be evenly distributed across the different levels of vaccine hesitancy,  Overall, 404 participants completed the survey. When we created a table looking at social economic status characteristics and vaccine hesitancy, only 93 participants answered all the corresponding questions. A majority of the participants did not answer the exempt reason question. After looking further into this, it may have been because participants were only given 3 options to choose from. There was no option for “other reason” for exemption. This may be an explanation to why the table includes only 25 percent of the participants. 

##### References:
1- Parental attitudes and perceptions associated with childhood vaccine exemptions in high-exemption schools. (n.d.). Retrieved November 29, 2020, from https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0198655

2- Benoit, S. L. (2020). The “Anti-Vax” Movement: A Quantitative Report on Vaccine Beliefs and Knowledge across Social Media [Data set]. Inter-university Consortium for Political and Social Research (ICPSR). https://doi.org/10.3886/E120505V1


